A Simple Solution of the Linearity Challenge

### 2025-12-28 paper/Examples.lagda.tex
tuning

\begin{remark}
    Writing processes in Agda would be easier if we could avoid using
    \AgdaFunction{curry∗}, namely if the constructors of the data type
    \AgdaDatatype{Proc} were naturally curried.
    %
    This is possible, but only if we are willing to sacrifice some clarity. To
    illustrate the problem with an example, below is the fully curried
    constructor \AgdaInductiveConstructor{fork} of a hypothetical data type
    \AgdaDatatype{Proc'}, obtained by expanding the definition of separating
    conjunction:
\begin{code}[hide]
data Proc' : Context → Set where
\end{code}
\begin{code}
  fork : ∀{A B Γ Δ Θ Θ₁ Θ₂} → Ch (A ⊗ B) Δ → Γ ≃ Δ + Θ →
         Proc' (A ∷ Θ₁) → Θ ≃ Θ₁ + Θ₂ → Proc' (B ∷ Θ₂) → Proc' (A ⊗ B ∷ Γ)
\end{code}

    This version of \AgdaInductiveConstructor{fork} is way less readable than
    the one we gave in \Cref{sec:process-agda} because of the (now visible)
    context splittings. We can recover some clarity and still obtain a curried
    version of \AgdaInductiveConstructor{fork} using the magic wand instead of
    Agda's native arrow:
\begin{code}[hide]
data Proc'' : Context → Set where
\end{code}
\begin{code}
  fork : ∀{A B} →
         ∀[ Ch (A ⊗ B) ⇒ (A ∷_) ⊢ Proc'' ─∗ (B ∷_) ⊢ Proc'' ─∗ Proc'' ]
\end{code}

    However, now we have to make a distinction between the first arrow
    \AgdaFunction{⇒} and the following magic wands to account for the
    appropriate amount of context splittings. We found this formulation of the
    constructors harder to explain and motivate in \Cref{sec:process-agda}.
    %
    Since none of the alternative definitions of \AgdaDatatype{Proc} was fully
    satisfactory, we preferred the cleanest and most readable version of the
    data type at the expense of additional clutter in this section.
    %
    \eoe
\end{remark}

>>>>

\begin{remark}
    Writing processes in Agda would be more pleasant if the constructors of the
    data type \AgdaDatatype{Proc} were naturally curried, instead of currying
    them on demand with \AgdaFunction{curry∗} as we do here.
    %
    Below is the naturally curried constructor \AgdaInductiveConstructor{fork}
    of a hypothetical data type \AgdaDatatype{Proc'}, obtained by expanding the
    definition of separating conjunction:
\begin{code}[hide]
data Proc' : Context → Set where
\end{code}
\begin{code}
  fork : ∀{A B Γ Δ Θ Θ₁ Θ₂} → Ch (A ⊗ B) Δ → Γ ≃ Δ + Θ →
         Proc' (A ∷ Θ₁) → Θ ≃ Θ₁ + Θ₂ → Proc' (B ∷ Θ₂) → Proc' (A ⊗ B ∷ Γ)
\end{code}

    This version of \AgdaInductiveConstructor{fork} is fully curried, but also
    less readable than the one we gave in \Cref{sec:process-agda} because of the
    (now visible) context splittings. We can recover some clarity and still
    obtain a curried version of \AgdaInductiveConstructor{fork} using
    (literally) the magic wand:
\begin{code}[hide]
data Proc'' : Context → Set where
\end{code}
\begin{code}
  fork : ∀{A B} →
         ∀[ Ch (A ⊗ B) ⇒ (A ∷_) ⊢ Proc'' ─∗ (B ∷_) ⊢ Proc'' ─∗ Proc'' ]
\end{code}

    However, the first arrow must be a plain implication \AgdaFunction{⇒} and
    not a magic wand to account for the appropriate amount of context
    splittings. We found this formulation of the constructors harder to explain
    and motivate in \Cref{sec:process-agda}.
    %
    Since none of the alternative definitions of \AgdaDatatype{Proc} was fully
    satisfactory, we preferred the most elegant version of the data type at the
    expense of additional clutter in this section.
    %
    \eoe
\end{remark}
